---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r}
library(dplyr)
library(lubridate)
activity_data <- read.csv("activity.csv") %>%
  # Convert integer intervals to time values so we can plot them properly.
  # TODO(laixer): Any easier way to do this?
  mutate(interval = origin + 60 * (60 * (interval %/% 100) + interval %% 100))
```

## What is mean total number of steps taken per day?

```{r}
library(dplyr)
library(ggplot2)
steps_by_date <- 
  activity_data %>% 
  filter(!is.na(steps)) %>% 
  group_by(date) %>% 
  summarize(total_steps = sum(steps))
ggplot(steps_by_date, aes(total_steps)) + geom_histogram(binwidth = 1000)
mean(steps_by_date$total_steps)
median(steps_by_date$total_steps)
```

## What is the average daily activity pattern?

```{r}
library(scales)
steps_by_interval <- 
  activity_data %>% 
  filter(!is.na(steps)) %>% 
  group_by(interval) %>% 
  summarize(mean_steps = mean(steps))
ggplot(steps_by_interval, aes(interval, mean_steps)) + 
  geom_point() + 
  geom_line() + 
  scale_x_datetime(labels = date_format("%H:%M"))
format(steps_by_interval$interval[which.max(steps_by_interval$mean_steps)], 
       format = "%H:%M")
```


## Imputing missing values

```{r}
sum(is.na(activity_data$steps))

activity_data_imputed <- activity_data %>% 
  mutate(steps = ifelse(is.na(steps), 
                        # TODO(laixer): Find less ugly way to do this.
                        steps_by_interval$mean_steps[steps_by_interval$interval == interval], 
                        steps))
steps_by_date_imputed <- 
  activity_data_imputed %>% 
  filter(!is.na(steps)) %>% 
  group_by(date) %>% 
  summarize(total_steps = sum(steps))
ggplot(steps_by_date_imputed, aes(total_steps)) + geom_histogram(binwidth = 1000)
mean(steps_by_date_imputed$total_steps)
median(steps_by_date_imputed$total_steps)
```

## Are there differences in activity patterns between weekdays and weekends?

```{r}
activity_data_imputed <- activity_data_imputed %>% 
  mutate(day_of_week=weekdays(ymd(date)), 
         type_of_day=as.factor(ifelse(day_of_week %in% c("Saturday", "Sunday"), "weekend", "weekday")))

steps_by_interval_and_type_of_day <- 
  activity_data_imputed %>% 
  filter(!is.na(steps)) %>% 
  group_by(interval, type_of_day) %>% 
  summarize(mean_steps = mean(steps))
ggplot(steps_by_interval_and_type_of_day, aes(interval, mean_steps)) + 
  facet_grid(type_of_day ~ .) +
  geom_point() + 
  geom_line() + 
  scale_x_datetime(labels = date_format("%H:%M"))
